Verification of concurrent design patterns with data

### 2019-02-13 2018-Arch-pNets/arch-pNets.tex
small typo
passe sur derniere annexe

@@

Parameterised Networks of synchronised automata (pNets) is a formalism for defining behavioural specification of distributed systems.
It inherited from the work of
Arnold on synchronisation vectors~\cite{Arnold1982}. 
In previous work~\cite{HMZ:PDP15}, it was shown pNets can be used to represent the behavioural semantics
of a system 
including value-passing and many kinds of synchronisation methods. We
used these results to give the semantics of various constructs and
languages for distributed objects, and to build a platform for design
and verification of distributed software components
\cite{CM:FMCO08,HKM-FASE16}. 
By nature, pNets is a  parametrised and hierarchical model.
Their structure is static, but unbounded, and this allows for model-checking
approaches even for reconfigurable applications.
Closed pNets were used to encode fully defined programs or systems,
while open pNets have ``holes'', playing the role of process
parameters. Such open systems can be used to define composition operators or structuring architectures.
It is possible to reason, in an SMT engine, on the symbolic automaton that represents the behaviour of a pNets with holes and that communicates value~\cite{HMZ-FORTE2016QBMZ-AVOCS18}. This encding of open pNets into Z3 that is under development is the starting point of this article. Our objective is to benefit from the possibility to reason on a pNet in an SMT engine in order to prove properties on BIP architectures.

@@

The standard pretty-print of OTs by the tool include the transitions
of the pLTSs (controllers). This is not a full ``proof-carrying
code'', but it allows for a better understanding of the construction
of each OT.

We first show the pNet input, then the twelve generated OTs in their
verbatim form.

\small\begin{verbatim}
FailureTimerArchi
--------------------
SV0 SVs: <tick(z, t, z1), fail(zone, b1), fail(b0), _>-->alpha0,
                          [((b1->b0)/\(z1=z/\(((b1?(zone)TopInt)))+t)))]
SV1 SVs: <cancel(b1), resume(b2), resume, _>-->alpha1, (b1=b2)
SV2 SVs: <timeout, timeout, _, ask>-->alpha2
SV3 SVs: <_, reset, _, reset>-->alpha3
SV4 SVs: <init, init, _, _>-->alpha4
pLTS: Timer
- Transition: t0---init-->t1, { t := 0 z := TopInt}
- Transition: t1---tick(z, t, z1)-->t1, [(t<GET_MAX z)], { t := (t+1) z := z1}
- Transition: t1---cancel(false)-->t1, {}
- Transition: t1---cancel(true)-->t1, { t := 0 z := TopInt}
- Transition: t1---timeout-->t1, [t in z], { t := (t+1)}
pLTS: Failure
- Transition: s0---init-->s1, { zone := minZone,maxZone}
- Transition: s1---timeout-->s1, {}
- Transition: s1---resume(false)-->s1, {}
- Transition: s1---fail(zone, false)-->s1, {}
- Transition: s1---fail(zone, true)-->s2, {}
- Transition: s2---fail(zone, false)-->s2, {}
- Transition: s2---resume(false)-->s2, {}
- Transition: s2---resume(true)-->s1, {}
- Transition: s2---timeout-->s3, {}
- Transition: s3---fail(zone, false)-->s3, {}
- Transition: s3---resume(false)-->s3, {}
- Transition: s3---reset-->s1, {}
Holes: B
Holes: E

       {t0---init-->t1, s0---init-->s1},   {},  [true],
       {t := 0, z := TopInt, zone := minZone,maxZone}
    OT1 ------------------------------------------------------------
        <t0_s0>-----init----><t1_s1>

       {t1---tick(z, t, z1)-->t1, s1---fail(zone, false)-->s1},   {---:hb_B-->},
       [(:hb_B=fail(b0:sva_SV0:1:1)/\(t<(GET_MAX z))
         /\((tick(z, t, z1))=(tick(z:sva_SV0:1:1, t:sva_SV0:1:1, z1:sva_SV0:1:1)))
         /\((fail(zone, false))=(fail(zone:sva_SV0:1:1, b1:sva_SV0:1:1)))
         /\(b1:sva_SV0:1:1=>b0:sva_SV0:1:1)
         /\(z1:sva_SV0:1:1=
         (z:sva_SV0:1:1/\((b1:sva_SV0:1:1?(zone:sva_SV0:1:1)TopInt))+t:sva_SV0:1:1)))],
       {t := t+1, z := z1}
    OT2 ------------------------------------------------------------
        <t1_s1>-----fail(b0:sva_SV0:1:1)----><t1_s1>

       {t1---tick(z, t, z1)-->t1, s1---fail(zone, true)-->s2},   {---:hb_B-->},
       [(:hb_B=fail(b0:sva_SV0:1:1)/\(t<(GET_MAX z))
         /\((tick(z,t, z1))=(tick(z:sva_SV0:1:1, t:sva_SV0:1:1, z1:sva_SV0:1:1)))
         /\((fail(zone, true))=(fail(zone:sva_SV0:1:1, b1:sva_SV0:1:1)))
         /\(b1:sva_SV0:1:1=>b0:sva_SV0:1:1)
         /\(z1:sva_SV0:1:1=
         (z:sva_SV0:1:1/\((b1:sva_SV0:1:1?(zone:sva_SV0:1:1)TopInt))+t:sva_SV0:1:1)))],
       {t := t+1, z := z1}
    OT3 ------------------------------------------------------------
        <t1_s1>-----fail(true)----><t1_s2>

       {t1---cancel(false)-->t1, s1---resume(false)-->s1},   {---:hb_B-->},
       [(:hb_B=fail(b0:sva_SV0:1:1)
         /\((cancel(false))=(cancel(b1:sva_SV1:1:2)))
         /\((resume(false))=(resume(b2:sva_SV1:1:2)))
         /\(b1:sva_SV1:1:2=b2:sva_SV1:1:2)],   {}
    OT4 ------------------------------------------------------------
        <t1_s1>-----resume----><t1_s1>

       {t1---timeout-->t1, s1---timeout-->s1},   {---:hb_E-->},
       [(:hb_E=ask)/\(t in z)],   {t := t+1}
    OT5 ------------------------------------------------------------
        <t1_s1>-----ask----><t1_s1>

       {t1---tick(z, t, z1)-->t1, s2---fail(zone, false)-->s2},   {---:hb_B-->},
       [(:hb_B=fail(b0:sva_SV0:1:1)/\(t<(GET_MAX z))
         /\((tick(z, t, z1))=(tick(z:sva_SV0:1:1, t:sva_SV0:1:1, z1:sva_SV0:1:1)))
         /\((fail(zone, false))=(fail(zone:sva_SV0:1:1, b1:sva_SV0:1:1)))
         /\(b1:sva_SV0:1:1=>b0:sva_SV0:1:1)
         /\(z1:sva_SV0:1:1=
         (z:sva_SV0:1:1/\((b1:sva_SV0:1:1?(zone:sva_SV0:1:1)TopInt))+t:sva_SV0:1:1)))],
       {t := t+1, z := z1}
    OT6 ------------------------------------------------------------
        <t1_s2>-----fail(b0:sva_SV0:1:1)----><t1_s2>

       {t1---cancel(false)-->t1, s2---resume(false)-->s2},   {---:hb_B-->},
       [(:hb_B=resume)/\((cancel(false))=(cancel(b1:sva_SV1:1:2)))
         /\((resume(false))=(resume(b2:sva_SV1:1:2)))/\(b1:sva_SV1:1:2=b2:sva_SV1:1:2)],   {}
    OT7 ------------------------------------------------------------
        <t1_s2>-----resume----><t1_s2>

       {t1---cancel(true)-->t1, s2---resume(true)-->s1},   {---:hb_B-->},
       [(:hb_B=resume)/\((cancel(true))=(cancel(b1:sva_SV1:1:2)))
         /\((resume(true))=(resume(b2:sva_SV1:1:2)))/\(b1:sva_SV1:1:2=b2:sva_SV1:1:2)],
       {t := 0, z := TopInt}
    OT8 ------------------------------------------------------------
        <t1_s2>-----resume----><t1_s1>

       {t1---timeout-->t1, s2---timeout-->s3},   {---:hb_E-->},
       [(:hb_E=ask)/\(t in z)],   {t := t+1}
    OT9 ------------------------------------------------------------
        <t1_s2>-----ask----><t1_s3>

       {t1---tick(z, t, z1)-->t1, s3---fail(zone, false)-->s3},   {---:hb_B-->},
       [(:hb_B=fail(b0:sva_SV0:1:1)/\(t<(GET_MAX z))
         /\((tick(z, t, z1))=(tick(z:sva_SV0:1:1, t:sva_SV0:1:1, z1:sva_SV0:1:1)))
         /\((fail(zone, false))=(fail(zone:sva_SV0:1:1, b1:sva_SV0:1:1)))
         /\(b1:sva_SV0:1:1=>b0:sva_SV0:1:1)
         /\(z1:sva_SV0:1:1=
         (z:sva_SV0:1:1/\((b1:sva_SV0:1:1?(zone:sva_SV0:1:1)TopInt))+t:sva_SV0:1:1)))],
       {t := t+1, z := z1}
    OT10 ------------------------------------------------------------
        <t1_s3>-----fail(b0:sva_SV0:1:1)----><t1_s3>

       {t1---cancel(false)-->t1, s3---resume(false)-->s3},   {---:hb_B-->},
       [(:hb_B=resume)/\((cancel(false))=(cancel(b1:sva_SV1:1:2)))
         /\((resume(false))=(resume(b2:sva_SV1:1:2)))/\(b1:sva_SV1:1:2=b2:sva_SV1:1:2)],   {}
    OT11 ------------------------------------------------------------
        <t1_s3>-----resume----><t1_s3>

        {s3---reset-->s1},   {---:hb_E-->},     [(:hb_E=ask)],   {}
    OT12 ------------------------------------------------------
        <t1_s3>-----reset----><t1_s1>

@@

OTs simplifiees a la main:
\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{TimerOADetailed}
  \caption{The open Automaton of the Failure Monitor Architecture}
  \label{schema:ArchFailure:OA}
\end{figure}

$  OT_1  = \openrule
{\emptyset, \{t:=0, z:=TopInt, zone:=[minZone,maxZone]\} }
{\ostate{00} \OTarrow{init} \ostate{11}}
$
\medskip

$  OT_2  = \openrule
{\{B\mapsto fail(b0)\},\\
  t<GetMax(z) \land b1=false \land b1=>b0 \land z1= (z \cap (b1?zone;TopInt)+t),
  \{z:=z1\} }
{\ostate{11} \OTarrow{fail} \ostate{11}}
$
\medskip

  $  OT_3  = \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=true \land b1=>b0 \land z1= (z \cap (b1?zone;TopInt)+t),
    \{z:=z1\} }
  {\ostate{11} \OTarrow{fail} \ostate{12}}
  $
  \medskip

  $  OT_4  = \openrule
  {\{B\mapsto resume\},
    b1=false \land b2=false \land b1=b2
    \{\}  }
  {\ostate{11} \OTarrow{resume} \ostate{11}}
  $
  \medskip

  $  OT_5  = \openrule
  {\{E\mapsto ask\},
    t \in z,
    \{t:=t+1\}  }
  {\ostate{11} \OTarrow{ask} \ostate{11}}
  $
  \medskip

  $  OT_6  = \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=false \land b1=>b0 \land z1=(z \cap (b1?zone;TopInt)+t),
    \{t:=t+1, z:=z1\} }
  {\ostate{12} \OTarrow{fail} \ostate{12}}
  $
  \medskip

  $  OT_7  = \openrule
  {\{B\mapsto resume\},
    b1=false \land b2=false \land b1=b2,
    \{\} }
  {\ostate{12} \OTarrow{resume} \ostate{12}}
  $
  \medskip

  $  OT_8  = \openrule
  {\{B\mapsto resume\},
    b1=true \land b2=true \land b1=b2, 
    \{t:=0, z:=TopInt\} }
  {\ostate{12} \OTarrow{resume} \ostate{11}}
  $
  \medskip

  $  OT_9  = \openrule
  {\{E\mapsto ask\},
    t \in z,
    \{t:=t+1\} }
  {\ostate{12} \OTarrow{ask} \ostate{13}}
  $
  \medskip

  $  OT_{10}  = \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=false \land b1=>b0 \land z1=(z \cap (b1?zone;TopInt)+t),
    \{t:=t+1, z:=z1\} }
  {\ostate{13} \OTarrow{fail} \ostate{13}}
  $
  \medskip

  $  OT_{11}  = \openrule
  {\{B\mapsto resume\},
    b1=false \land b2=false \land b1=b2,
    \{\} }
  {\ostate{13} \OTarrow{resume} \ostate{13}}
  $
  \medskip

  $  OT_{12}  = \openrule
  {\{E\mapsto reset\},
    true,
    \{\} }
  {\ostate{13} \OTarrow{reset} \ostate{11}}
  $

  Now the version with the Max Progress encoding.
 This modifies both the pLTS transitions, and the synchronisation vectors:

\small\begin{verbatim}
FailureTimerArchi
--------------------
SV0 SVs: <tick(z, t, z1), fail(zone, b1), fail(b0), _>-->alpha0,
                          [((b1->b0)/\(z1=z/\(((b1?(zone)TopInt)))+t)))]
SV1 SVs: <cancel(b1), resume(b2), resume, _>-->alpha1, (b1=b2)
SV2 SVs: <timeout, timeout, _, ask>-->alpha2
SV3 SVs: <_, reset, _, reset>-->alpha3
SV4 SVs: <init, init, _, _>-->alpha4
pLTS: Timer
- Transition: t0---init-->t1, { t := 0 z := TopInt}
- Transition: t1---tick(z, t, z1)-->t1, [(t<GET_MAX z)], { t := (t+1) z := z1}
- Transition: t1---cancel(false)-->t1, {}
- Transition: t1---cancel(true)-->t1, { t := 0 z := TopInt}
- Transition: t1---timeout-->t1, [t in z], { t := (t+1)}
pLTS: Failure
- Transition: s0---init-->s1, { zone := minZone,maxZone}
- Transition: s1---timeout-->s1, {}
- Transition: s1---resume(false)-->s1, {}
- Transition: s1---fail(zone, false)-->s1, {}
- Transition: s1---fail(zone, true)-->s2, {}
- Transition: s2---fail(zone, false)-->s2, {}
- Transition: s2---resume(false)-->s2, {}
- Transition: s2---resume(true)-->s1, {}
- Transition: s2---timeout-->s3, {}
- Transition: s3---fail(zone, false)-->s3, {}
- Transition: s3---resume(false)-->s3, {}
- Transition: s3---reset-->s1, {}
Holes: B
Holes: E
\end{verbatim}

  The new (simplied) OA is in figure \ref{schema:ArchFailure:OA-MaxProgress}.
  Here we give the detail only of one the modifed OTs:

@@

\small\begin{verbatim}
	    {t1---tick(z, t, z1)-->t1, s1---fail(zone, true, false)-->s1},
            {B|->fail(b0:sva_SV0:1:1)},
            [(t<(GET_MAX z))/\((tick(z, t, z1))=(tick(z:sva_SV0:1:1, t:sva_SV0:1:1, z1:sva_SV0:1:1)))
              /\((fail(zone, true, false))=(fail(zone:sva_SV0:1:1, bf:sva_SV0:1:1, b1:sva_SV0:1:1)))
              /\(b1:sva_SV0:1:1=>b0:sva_SV0:1:1)
              /\((bf:sva_SV0:1:1/\b0:sva_SV0:1:1)=>b1:sva_SV0:1:1)
              /\(z1:sva_SV0:1:1=(z:sva_SV0:1:1 |Intersection| ((b1:sva_SV0:1:1?(zone:sva_SV0:1:1):(TopInt))+t:sva_SV0:1:1)))],
            {t := t+1, z := z1}
	OT:2 ------------------------------------------------------------
		<t1_s1>-----fail(false)----><t1_s1>

\end{verbatim}

>>>>

% %%%%%%%%%%%%%%%%%%%%%
\makeatletter
\newcommand{\verbatimfont}[1]{\renewcommand{\verbatim@font}{\ttfamily#1}}
\makeatother
% %%%%%%%%%%%%%%%%%%%%%

@@

Parameterised Networks of synchronised automata (pNets) is a formalism for defining behavioural specification of distributed systems.
It inherited from the work of
Arnold on synchronisation vectors~\cite{Arnold1982}. 
In previous work~\cite{HMZ:PDP15}, it was shown pNets can be used to represent the behavioural semantics
of a system 
including value-passing and many kinds of synchronisation methods. We
used these results to give the semantics of various constructs and
languages for distributed objects, and to build a platform for design
and verification of distributed software components
\cite{CM:FMCO08,HKM-FASE16}. 
By nature, pNets is a  parametrised and hierarchical model.
Their structure is static, but unbounded, and this allows for model-checking
approaches even for reconfigurable applications.
Closed pNets were used to encode fully defined programs or systems,
while open pNets have ``holes'', playing the role of process
parameters. Such open systems can be used to define composition operators or structuring architectures.
It is possible to reason, in an SMT engine, on the symbolic automaton that represents the behaviour of a pNets with holes and that communicates value~\cite{HMZ-FORTE2016QBMZ-AVOCS18}. This encoding of open pNets into Z3 that is under development is the starting point of this article. Our objective is to benefit from the possibility to reason on a pNet in an SMT engine in order to prove properties on BIP architectures.

@@

We first show the input pNet, with its 5 synchronisation vectors, and 2 controllers.

\noindent
\verbatimfont{\small}
\begin{verbatim}
SV0: <tick(z, t, z1), fail(zone, b1), fail(b0), _> --> fail(b0),
         [(b1->b0)/\(z1=z |Intersection| ((b1?zone:TopInt)+t))]
SV1: <cancel(b1), resume(b2), resume, _> --> resume, (b1=b2)
SV2: <timeout, timeout, _, ask> --> ask
SV3: <_, reset, _, reset> --> reset
SV4: <init, init, _, _> --> init
pLTS: Timer
- Transition: t0---init-->t1, { t := 0 z := TopInt}
- Transition: t1---tick(z, t, z1)-->t1, [(t<GET_MAX z)], { t := (t+1) z := z1}
- Transition: t1---cancel(false)-->t1, {}
- Transition: t1---cancel(true)-->t1, { t := 0 z := TopInt}
- Transition: t1---timeout-->t1, [t in z], { t := (t+1)}
pLTS: Control
- Transition: s0---init-->s1, { zone := minZone,maxZone}
- Transition: s1---timeout-->s1, {}
- Transition: s1---resume(false)-->s1, {}
- Transition: s1---fail(zone, false)-->s1, {}
- Transition: s1---fail(zone, true)-->s2, {}
- Transition: s2---fail(zone, false)-->s2, {}
- Transition: s2---resume(false)-->s2, {}
- Transition: s2---resume(true)-->s1, {}
- Transition: s2---timeout-->s3, {}
- Transition: s3---fail(zone, false)-->s3, {}
- Transition: s3---resume(false)-->s3, {}
- Transition: s3---reset-->s1, {}
Holes: B
Holes: E
\end{verbatim}

Then some of the twelve generated OTs in their verbatim form. In this format
the first element of the premisse is a list of the moves of the pLTS
participating in the transition. This is not part of the OT, but
provides a trace of its construction. Also, the variables occuring in
the actions and predicate expressions have names helping to understand
where they come from (e.g. \verb|b0:sva_SV0| in $OT2$ below).
This is not a full ``proof-carrying
code'', but it allows for a better understanding of the construction
of each OT.

\verbatimfont{\small}
\begin{verbatim}
   {t0---init-->t1, s0---init-->s1},   {},  [true],
   {t := 0, z := TopInt, zone := minZone,maxZone}
OT1 ------------------------------------------------------------
    <t0_s0>-----init----><t1_s1>

   {t1---tick(z, t, z1)-->t1, s1---fail(zone, false)-->s1},   {---:hb_B-->},
   [(:hb_B=fail(b0:sva_SV0)/\(t<(GET_MAX z))
     /\((tick(z, t, z1))=(tick(z:sva_SV0, t:sva_SV0, z1:sva_SV0)))
     /\((fail(zone, false))=(fail(zone:sva_SV0, b1:sva_SV0)))
     /\(b1:sva_SV0=>b0:sva_SV0)
     /\(z1:sva_SV0=
     (z:sva_SV0  |Intersection| ((b1:sva_SV0?(zone:sva_SV0)TopInt))+t:sva_SV0)))],
   {t := t+1, z := z1}
OT2 ------------------------------------------------------------
    <t1_s1>-----fail(b0:sva_SV0)----><t1_s1>

   {t1---tick(z, t, z1)-->t1, s1---fail(zone, true)-->s2},   {---:hb_B-->},
   [(:hb_B=fail(b0:sva_SV0)/\(t<(GET_MAX z))
     /\((tick(z,t, z1))=(tick(z:sva_SV0, t:sva_SV0, z1:sva_SV0)))
     /\((fail(zone, true))=(fail(zone:sva_SV0, b1:sva_SV0)))
     /\(b1:sva_SV0=>b0:sva_SV0)
     /\(z1:sva_SV0=
     (z:sva_SV0  |Intersection| ((b1:sva_SV0?(zone:sva_SV0)TopInt))+t:sva_SV0)))],
   {t := t+1, z := z1}
OT3 ------------------------------------------------------------
    <t1_s1>-----fail(true)----><t1_s2>

   {t1---cancel(false)-->t1, s1---resume(false)-->s1},   {---:hb_B-->},
   [(:hb_B=fail(b0:sva_SV0)
     /\((cancel(false))=(cancel(b1:sva_SV1:1:2)))
     /\((resume(false))=(resume(b2:sva_SV1:1:2)))
     /\(b1:sva_SV1:1:2=b2:sva_SV1:1:2)],   {}
OT4 ------------------------------------------------------------
    <t1_s1>-----resume----><t1_s1>

@@

%%    {t1---timeout-->t1, s1---timeout-->s1},   {---:hb_E-->},
%%    [(:hb_E=ask)/\(t in z)],   {t := t+1}
%% OT5 ------------------------------------------------------------
%%     <t1_s1>-----ask----><t1_s1>

%%    {t1---tick(z, t, z1)-->t1, s2---fail(zone, false)-->s2},   {---:hb_B-->},
%%    [(:hb_B=fail(b0:sva_SV0)/\(t<(GET_MAX z))
%%      /\((tick(z, t, z1))=(tick(z:sva_SV0, t:sva_SV0, z1:sva_SV0)))
%%      /\((fail(zone, false))=(fail(zone:sva_SV0, b1:sva_SV0)))
%%      /\(b1:sva_SV0=>b0:sva_SV0)
%%      /\(z1:sva_SV0=
%%      (z:sva_SV0  |Intersection| ((b1:sva_SV0?(zone:sva_SV0)TopInt))+t:sva_SV0)))],
%%    {t := t+1, z := z1}
%% OT6 ------------------------------------------------------------
%%     <t1_s2>-----fail(b0:sva_SV0)----><t1_s2>

%%    {t1---cancel(false)-->t1, s2---resume(false)-->s2},   {---:hb_B-->},
%%    [(:hb_B=resume)/\((cancel(false))=(cancel(b1:sva_SV1:1:2)))
%%      /\((resume(false))=(resume(b2:sva_SV1:1:2)))/\(b1:sva_SV1:1:2=b2:sva_SV1:1:2)],
%%    {}
%% OT7 ------------------------------------------------------------
%%     <t1_s2>-----resume----><t1_s2>

%%    {t1---cancel(true)-->t1, s2---resume(true)-->s1},   {---:hb_B-->},
%%    [(:hb_B=resume)/\((cancel(true))=(cancel(b1:sva_SV1:1:2)))
%%      /\((resume(true))=(resume(b2:sva_SV1:1:2)))/\(b1:sva_SV1:1:2=b2:sva_SV1:1:2)],
%%    {t := 0, z := TopInt}
%% OT8 ------------------------------------------------------------
%%     <t1_s2>-----resume----><t1_s1>

%%    {t1---timeout-->t1, s2---timeout-->s3},   {---:hb_E-->},
%%    [(:hb_E=ask)/\(t in z)],   {t := t+1}
%% OT9 ------------------------------------------------------------
%%     <t1_s2>-----ask----><t1_s3>

%%    {t1---tick(z, t, z1)-->t1, s3---fail(zone, false)-->s3},   {---:hb_B-->},
%%    [(:hb_B=fail(b0:sva_SV0)/\(t<(GET_MAX z))
%%      /\((tick(z, t, z1))=(tick(z:sva_SV0, t:sva_SV0, z1:sva_SV0)))
%%      /\((fail(zone, false))=(fail(zone:sva_SV0, b1:sva_SV0)))
%%      /\(b1:sva_SV0=>b0:sva_SV0)
%%      /\(z1:sva_SV0=
%%      (z:sva_SV0  |Intersection| ((b1:sva_SV0?(zone:sva_SV0)TopInt))+t:sva_SV0)))],
%%    {t := t+1, z := z1}
%% OT10 ------------------------------------------------------------
%%     <t1_s3>-----fail(b0:sva_SV0)----><t1_s3>

%%    {t1---cancel(false)-->t1, s3---resume(false)-->s3},   {---:hb_B-->},
%%    [(:hb_B=resume)/\((cancel(false))=(cancel(b1:sva_SV1:1:2)))
%%      /\((resume(false))=(resume(b2:sva_SV1:1:2)))/\(b1:sva_SV1:1:2=b2:sva_SV1:1:2)],   {}
%% OT11 ------------------------------------------------------------
%%     <t1_s3>-----resume----><t1_s3>

%%     {s3---reset-->s1},   {---:hb_E-->},     [(:hb_E=ask)],   {}
%% OT12 ------------------------------------------------------
%%     <t1_s3>-----reset----><t1_s1>

Now we list the full set of OTs, in latex form, for better
readability. This comes as a complement to the open automaton in
Fig. \ref{schema:ArchFailure:OA}, in which we only have a partial view
of the OTs details.

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{TimerOAFullDetailed}
  \caption{The open Automaton of the Failure Monitor Architecture}
  \label{schema:ArchFailure:OA}
\end{figure}

\medskip
\noindent
$  OT_1  = \openrule
{\emptyset, \{t:=0, z:=TopInt, zone:=[minZone,maxZone]\} }
{\ostate{00} \OTarrow{init} \ostate{11}}
$

\medskip    
\noindent
$  OT_2  = \openrule
{\{B\mapsto fail(b0)\},\\
  t<GetMax(z) \land b1=false \land b1=>b0 \land z1= (z \cap (b1?zone;TopInt)+t),\\
  \{z:=z1\} }
{\ostate{11} \OTarrow{fail} \ostate{11}}
$
\medskip

\noindent
  $  OT_3  = \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=true \land b1=>b0 \land z1= (z \cap (b1?zone;TopInt)+t),\\
    \{z:=z1\} }
  {\ostate{11} \OTarrow{fail} \ostate{12}}
  $
  \medskip

\noindent
  $  OT_4  = \openrule
  {\{B\mapsto resume\},
    b1=false \land b2=false \land b1=b2
    \{\}  }
  {\ostate{11} \OTarrow{resume} \ostate{11}}
  $
  \medskip

\noindent
  $  OT_5  = \openrule
  {\{E\mapsto ask\},
    t \in z,
    \{t:=t+1\}  }
  {\ostate{11} \OTarrow{ask} \ostate{11}}
  $
  \medskip

\noindent
  $  OT_6  = \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=false \land b1=>b0 \land z1=(z \cap (b1?zone;TopInt)+t),\\
    \{t:=t+1, z:=z1\} }
  {\ostate{12} \OTarrow{fail} \ostate{12}}
  $
  \medskip

\noindent
  $  OT_7  = \openrule
  {\{B\mapsto resume\},
    b1=false \land b2=false \land b1=b2,
    \{\} }
  {\ostate{12} \OTarrow{resume} \ostate{12}}
  $
  \medskip

\noindent
  $  OT_8  = \openrule
  {\{B\mapsto resume\},
    b1=true \land b2=true \land b1=b2, 
    \{t:=0, z:=TopInt\} }
  {\ostate{12} \OTarrow{resume} \ostate{11}}
  $
  \medskip

\noindent
  $  OT_9  = \openrule
  {\{E\mapsto ask\},
    t \in z,
    \{t:=t+1\} }
  {\ostate{12} \OTarrow{ask} \ostate{13}}
  $
  \medskip

\noindent
  $  OT_{10}  = \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=false \land b1=>b0 \land z1=(z \cap (b1?zone;TopInt)+t),\\
    \{t:=t+1, z:=z1\} }
  {\ostate{13} \OTarrow{fail} \ostate{13}}
  $
  \medskip

\noindent
  $  OT_{11}  = \openrule
  {\{B\mapsto resume\},
    b1=false \land b2=false \land b1=b2,
    \{\} }
  {\ostate{13} \OTarrow{resume} \ostate{13}}
  $
  \medskip

\noindent
  $  OT_{12}  = \openrule
  {\{E\mapsto reset\},
    true,
    \{\} }
  {\ostate{13} \OTarrow{reset} \ostate{11}}
  $

  \medskip

  \subsubsection{Properties}

  In Fig. \ref{schema:ArchFailure:OA}, you can see that some
  properties from appendix \app{properties} do not hold.

  In particular, the first safety property:
  \begin{equation}*
    \label{eq:prop:reset:can}
    \AG \bigl(\PortFail(true) \rightarrow \EU[]{\lnot \PortFail(true)}{(\PortResume(true) \lor \PortReset)}\bigr).
  \end{equation}
is not true, because the $OT_2$ loop on state $\ostate{11}$ stays in
state 11 where you need another $\PortFail(true)$ before enabling a
resume (or reset). This is because without priorities we can have an
activity on the dandling ``fail'' port without synchronising with the
others, even if they are available.

This can be sloved by the version with the Max Progress encoding.
This modifies both the pLTS transitions, and the synchronisation
vectors, as shown below:

\small\begin{verbatim}
SV0: <tick(z, t, z1), fail(zone, bf, b1), fail(b0), _> --> fail(b0),
         [(b1=>b0)/\((bf/\b0)=>b1)/\(z1=(z |Intersection| ((b1?zone:TopInt)+t)))]
SV1: <cancel(bc, b1), resume(br, b2), resume, _>-->resume, (b1=b2)/\((bc/\br)=>(b1/\b2))
SV2: <timeout, timeout, _, ask> --> ask
SV3: <_, reset, _, reset> --> reset
SV4: <init, init, _, _> --> init

pLTS: Timer
- Transition: t0---init-->t1, { t := 0 z := TopInt}
- Transition: t1---tick(z, t, z1)-->t1, [t<(GET_MAX z)], { t := t+1 z := z1}
- Transition: t1---cancel(true, false)-->t1, {}
- Transition: t1---cancel(true, true)-->t1, { t := 0 z := TopInt}
- Transition: t1---timeout-->t1, [t in z], { t := 0 z := TopInt}
pLTS: Failure
- Transition: s0---init-->s1, { zone := minZone,maxZone}
- Transition: s1---timeout-->s1, {}
- Transition: s1---resume(false, false)-->s1, {}
- Transition: s1---fail(zone, true, false)-->s1, {}
- Transition: s1---fail(zone, true, true)-->s2, {}
- Transition: s2---fail(zone, false, false)-->s2, {}
- Transition: s2---resume(true, false)-->s2, {}
- Transition: s2---resume(true, true)-->s1, {}
- Transition: s2---timeout-->s3, {}
- Transition: s3---fail(zone, false, false)-->s3, {}
- Transition: s3---resume(false, false)-->s3, {}
- Transition: s3---reset-->s1, {}
\end{verbatim}

  The new OA is shown in figure \ref{schema:ArchFailure:OA-MaxProgress}.

@@

  Here we give the detail of the modification of $OT_2$:

\small\begin{verbatim}
	    {t1---tick(z, t, z1)-->t1, s1---fail(zone, true, false)-->s1},
            {B|->fail(b0:sva_SV0)},
            [(t<(GET_MAX z))/\((tick(z, t, z1))=(tick(z:sva_SV0, t:sva_SV0, z1:sva_SV0)))
              /\((fail(zone, true, false))=(fail(zone:sva_SV0, bf:sva_SV0, b1:sva_SV0)))
              /\(b1:sva_SV0=>b0:sva_SV0)
              /\((bf:sva_SV0/\b0:sva_SV0)=>b1:sva_SV0)
              /\(z1:sva_SV0=(z:sva_SV0 |Intersection| ((b1:sva_SV0?(zone:sva_SV0):(TopInt))+t:sva_SV0)))],
            {t := t+1, z := z1}
	OT:2 ------------------------------------------------------------
		<t1_s1>-----fail(b0:sva_SV0)----><t1_s1>
\end{verbatim}

As \verb|bf| is true, and \verb|b1| is false, the predicate fragment
\verb|((bf/\b0)=>b1)| imposes that \verb|b0| is false.

\subsubsection{Scaling-up: composing architectures}

\todoEM{Il faudrait (mais a-t-on le temps?): faire un dessin de la
  composition au niveau BIP, et au niveau pNet. Un resultat brut, vaguement
  correct, montre 36 OTs, modulo les petites questions de mon mail de
  mercredi midi...
  Au moins dessiner cet automate, et ecrire les proprietes qu'il
  verifie... }

Primary version of the composed pNet:

\begin{verbatim}
SV0: <tick(z, t, z1), fail(zone1, b11), fail(zone2, b12), fail1(b01), fail2(b02), _> --> fail(b01, b02), 
     (z1 = z |Intersection| ((b11?zone1:TopInt)+t) |Intersection| ((b12?zone2:TopInt)+t))
     /\(b11=>b01)/\(b12=>b02)
SV1: <cancel(b1), resume(b21), resume(b22), resume1, resume2, _>-->resume, (b1=b21)/\(b1=b22)
SV2: <timeout, timeout, timeout, _, _, ask>-->ask
SV3: <_, reset, reset, _, _, reset>-->reset
SV4: <init, init, init, _, _, _>-->init

pLTS: Timer
- Transition: t0---init-->t1, { t := 0 z := TopInt}
- Transition: t1---tick(z, t, z1)-->t1, [t<(GET_MAX z)], { t := t+1 z := z1}
- Transition: t1---cancel(false)-->t1, {}
- Transition: t1---cancel(true)-->t1, { t := 0 z := TopInt}
- Transition: t1---timeout-->t1, [t in z], { t := 0 z := TopInt}

pLTS: FC1
- Transition: s0---init-->s1, { zone1 := minZone1,maxZone1}
- Transition: s1---timeout-->s1, {}
- Transition: s1---resume(false)-->s1, {}
- Transition: s1---fail(zone1, false)-->s1, {}
- Transition: s1---fail(zone1, true)-->s2, {}
- Transition: s2---fail(zone1, false)-->s2, {}
- Transition: s2---resume(false)-->s2, {}
- Transition: s2---resume(true)-->s1, {}
- Transition: s2---timeout-->s3, {}
- Transition: s3---fail(zone1, false)-->s3, {}
- Transition: s3---timeout-->s3, {}
- Transition: s3---resume(false)-->s3, {}
- Transition: s3---reset-->s1, {}

pLTS: FC2
- Transition: s0---init-->s1, { zone2 := minZone2,maxZone2}
- Transition: s1---timeout-->s1, {}
- Transition: s1---resume(false)-->s1, {}
- Transition: s1---fail(zone2, false)-->s1, {}
- Transition: s1---fail(zone2, true)-->s2, {}
- Transition: s2---fail(zone2, false)-->s2, {}
- Transition: s2---resume(false)-->s2, {}
- Transition: s2---resume(true)-->s1, {}
- Transition: s2---timeout-->s3, {}
- Transition: s3---fail(zone2, false)-->s3, {}
- Transition: s3---timeout-->s3, {}
- Transition: s3---resume(false)-->s3, {}
- Transition: s3---reset-->s1, {}
